|

### 2025-01-29 doc/paper.qmd
paper upd

# List of Abbreviations
\setlength{\parindent}{0em}
**BHR**: Buy-and-Hold Returns   
**CRSP**: Center for Research in Security Prices  
**CQA**: Corporate Decision-Making and Quantitative Analysis  
**DV**: Dependent Variable  
**HU**: Humboldt-Universität zu Berlin  
**IDE**: Integrated Development Environment  
**IV**: Independent Variable  
**TREAT**: TRR 266 Template for Reproducible Empirical Accounting Research  
**US**: United States  
**WRDS**: Wharton Research Data Services

@@

Building on these foundations, Assignment III investigates the informativeness of quarterly earnings announcements and their contribution to annual share price movements. The project replicates and extends @ball_how_2008, assessing the extent to which earnings releases serve as a source of new information for financial markets. By integrating multiple datasets from WRDS and applying an event-study methodology, this study evaluates the economic role of earnings disclosures in market efficiency, showcasing the valuation purpose of accounting. At the same time, earnings reports may play a crucial role in contracting by influencing managerial compensation, debt agreements, and other contractual mechanisms tied to financial performance [@ball_how_2008]. The integration of multiple databases adds complexity, requiring a detailed understanding of their structures and careful scripting to extract, align, and analyze data effectively. Key tasks include replicating and comparing original results, extending the analysis to data from 2007–2023, and applying the methodology to a non-U.S. country to explore cross-market variations.

@@

The paper is structured into sections corresponding to Tasks 1–3 (@sec-task1, @sec-task2, @sec-task3), each detailing the research design choices and assumptions, documenting the replication steps, and analyzing the replication results for the respective segment. @sec-task1 follows an exact replication approach by replicating key tables and figures from the original study using US data (1972–2006) to establish baseline results. @sec-task2 extends the analysis through empirical generalization, applying the same methodology to 2007–2023 US data to compare how results evolve over time. @sec-task3 employs a generalization and extension approach by applying the methodology to a non-US country, requiring adjustments for different databases and market structures. This framework aligns with the replication taxonomy outlined by @salterio_replication_2022. Finally, @sec-task4 sketches the survey design as a non-archival study that allows evaluating the key findings of the seminal paper, following Bloomfield’s (2016) triangulation principle. The concluding remarks are provided in @sec-conclusion.  The entire Python computation code for return calculations is available in `code/python/do_analysis.py` for detailed review if necessary. This paper focuses on the replication process, presenting visualizations, and discussing the results.

@@

@@

## Replication Steps {#sec-replication_steps1}
The analysis involves gathering and filtering stock return and earnings announcement data, linking CRSP stock data with Compustat fundamentals, and estimating the relationship between annual returns and earnings-announcement window returns. The replication process follows a structured workflow of data pull, preparation, and analysis, ensuring consistency with the original study.

@@

@@

## Results {#sec-results1}
Comparing the replication results of Figure 1 (see @fig-market-shares) with those from @ball_how_2008 (see Figure 1 below) reveals both similarities and discrepancies.

```{python}
#| label: fig1_original_png
#| fig-cap: "Original Figure 1 from Ball and Shivakumar (2008)"
#| echo: false

from IPython.display import Image

Image(filename="fig1_original.png")
```

@fig-market-shares illustrates the... Countries are sorted in a specific order reflecting the replication of Figure 3 from the European Commission’s 2024 report. 
```{python}
#| label: fig-market-shares
#| fig-cap: "Replicated Figure 3: Audit Firms' Market Shares in the EU for 2021"
#| fig-align: center
#| echo: false

import pickle
import matplotlib.pyplot as plt

# Load the figure from the pickle file
with open('../output/figure3_market_shares.pickle', 'rb') as f:
    fig = pickle.load(f)

# Render the figure explicitly
plt.figure(fig.number)
plt.show()
```

@@

In summary, the replication effectively captures the original report's key findings, confirming the dominance of the Big 4 and the broader 10KAP group in the European audit market. The analysis highlights varying degrees of market concentration across countries, with some demonstrating a more decentralized audit landscape. While discrepancies in individual country results, such as Romania, underline the importance of methodological alignment, the strong similarity between the overall market shares reported in the original report and the replicated results underscores the robustness and reliability of the findings.
##we confirm? "The results are consistent with the view that the primary economic role of reported earnings is not to provide timely new information to the share market. By inference, that role lies elsewhere, for example, in settling debt and compensation contracts and in disciplining prior information,including more timely managerial disclosures of information originating in the firm’s accounting system. The comparatively low “surprise” content of earnings announcements is to be expected, for a variety of reasons. By its nature, accounting earnings is low frequency (quarterly), not discretionary (announced every quarter, independent of whether there is substantial new information to report), and primarily backward-looking. Other information, and hence revision in share price, is comparatively high frequency, frequently discretionary (released only when there is substantial information to report), and both forward-looking and backward-looking."

# Task 2 - Extending the Analysis to 2007–2023 {#sec-task2}

This Section extends the analysis to 2007–2023, a period defined by major economic events such as the 2008 financial crisis and the COVID-19 pandemic, which may have influenced the informativeness of earnings announcements. While the task is relatively straightforward in terms of programming—requiring adjustments to the sample period—it demands careful economic interpretation of observed trends, market shifts, and potential structural changes that may influence the relationship between earnings announcements and stock returns.

@@

Assumptions from @sec-research_design_assumptions1 still hold

@@

# Task 3 - Cross-Country Replication {#sec-task3}

Task 3 extends the analysis beyond the United States by replicating Figure 1 using data from a non-U.S. market, providing insights into how earnings informativeness varies across different regulatory environments and market structures. This task requires mapping variables from CRSP/Compustat to their equivalents in Worldscope/Datastream, adjusting for differences in accounting standards, market liquidity, and institutional factors. While the core replication methodology remains the same, careful data handling and economic interpretation are crucial to account for cross-country differences.

## Research Design Choices and Assumptions {#sec-research_design_assumptions3}

It is important to use same logic in defining event windows as in Task 1 to ensure consistency https://community.developers.refinitiv.com/discussion/106705/how-to-get-only-trading-days-in-datastream-data/p1?tab=accepted
country-specific market calendars ?

We need to check whether non-US databases (Worldscope/Datastream) use the same announcement date convention or if they report earnings by fiscal period.

## Replication Steps {#sec-replication_steps3}
## Results {#sec-results3}

@@

The study contributes to the ongoing discussion on market efficiency and the economic role of corporate disclosures by providing empirical evidence on how earnings informativeness has evolved over time and across jurisdictions. The cross-country analysis highlights variations in market responses to earnings announcements, suggesting that institutional factors, market structures, and investor behavior play a role in shaping how financial information is processed. These findings have implications for investors, analysts, and policymakers, as they underscore the importance of regulatory consistency and market transparency in shaping the effectiveness of financial reporting.

>>>>

# List of Abbreviations
\setlength{\parindent}{0em}
**BHR**: Buy-and-Hold Returns   
**CRSP**: Center for Research in Security Prices  
**CQA**: Corporate Decision-Making and Quantitative Analysis  
**DS**: Datastream  
**DV**: Dependent Variable  
**HU**: Humboldt-Universität zu Berlin  
**IDE**: Integrated Development Environment  
**IV**: Independent Variable  
**TREAT**: TRR 266 Template for Reproducible Empirical Accounting Research  
**US**: United States  
**WRDS**: Wharton Research Data Services  
**WSCP**: Worldscope

@@

Building on these foundations, Assignment III investigates the informativeness of quarterly earnings announcements and their contribution to annual share price movements. The project replicates and extends @ball_how_2008, assessing the extent to which earnings releases serve as a source of new information for financial markets. By integrating multiple datasets from WRDS and applying an event-study methodology, this study evaluates the economic role of earnings disclosures in market efficiency, showcasing the valuation purpose of accounting. At the same time, earnings reports may play a crucial role in contracting by influencing managerial compensation, debt agreements, and other contractual mechanisms tied to financial performance [@ball_how_2008]. The integration of multiple databases adds complexity, requiring a detailed understanding of their structures and careful scripting to extract, align, and analyze data effectively. Key tasks include replicating and comparing original results, extending the analysis to data from 2007–2023, and applying the methodology to Canada as a non-U.S. country to explore cross-market variations.

@@

The paper is structured into sections corresponding to Tasks 1–3 (@sec-task1, @sec-task2, @sec-task3), each detailing the research design choices and assumptions, documenting the replication steps, and analyzing the replication results for the respective segment. @sec-task1 follows an exact replication approach by replicating key tables and figures from the original study using US data (1972–2006) to establish baseline results. @sec-task2 extends the analysis through empirical generalization, applying the same methodology to 2007–2023 US data to compare how results evolve over time. @sec-task3 employs a generalization and extension approach by applying the methodology to a non-US country, requiring adjustments for different databases and market structures. This framework aligns with the replication taxonomy outlined by @salterio_replication_2022. Finally, @sec-task4 sketches the survey design as a non-archival study that allows evaluating the key findings of the seminal paper, following Bloomfield’s (2016) triangulation principle. The concluding remarks are provided in @sec-conclusion.  The entire Python computation code for regression run is available in `code/python/do_analysis.py` and `code/python/do_analysis-wscp.py` for detailed review if necessary. This paper focuses on the replication process, presenting visualizations, and discussing the results.

@@

7. Same as @ball_how_2008, I group earnings announcements based on their actual announcement dates (`RDQ` from Compustat) rather than the fiscal periods they cover. Since Compustat provides the precise earnings announcement date and CRSP aligns stock returns with trading days (`date` in CRSP), this approach ensures consistency. So, no adjustments are made to realign quarters.

@@

## Replication Steps {#sec-replication_steps1}
The analysis involves gathering and filtering stock return and earnings announcement data, linking CRSP stock data with Compustat firm-level accounting data fundamentals, and estimating the relationship between annual returns and earnings-announcement window returns. The replication process follows a structured workflow of data pull, preparation, and analysis, ensuring consistency with the original study.

@@

Since @ball_how_2008 do not explicitly describe their return computation, I compute Buy-and-Hold Returns (BHR) following the cumulative return formula as discussed in @gundersen_returns_2022. The BHR formula captures compounded returns over multiple days, reflecting the stock’s actual performance.

@@

## Results {#sec-results1}
Comparing the replication results of Figure 1 (see @fig-market-shares) with those from @ball_how_2008 (see @fig-fig1_original) reveals both similarities and discrepancies.

```{python}
#| label: fig-fig1_original
#| fig-cap: "Original Figure 1 from Ball and Shivakumar (2008)"
#| echo: false

import matplotlib.pyplot as plt
import matplotlib.image as mpimg

img = mpimg.imread("fig1_original.png")
plt.figure(figsize=(12, 20))
plt.imshow(img)
plt.axis("off")
plt.show()

# used in AssI Solution, I use knitr for availability of cross-referencing.
#| label: fig1_original_png
#| fig-cap: "Original Figure 1 from Ball and Shivakumar (2008)"
#| echo: false

#from IPython.display import Image

#Image(filename="fig1_original.png")
#
```

@fig-market-shares illustrates the... Countries are sorted in a specific order reflecting the replication of Figure 3 from the European Commission’s 2024 report.

@@

In summary, the replication effectively captures the original report's key findings, confirming the dominance of the Big 4 and the broader 10KAP group in the European audit market. The analysis highlights varying degrees of market concentration across countries, with some demonstrating a more decentralized audit landscape. While discrepancies in individual country results, such as Romania, underline the importance of methodological alignment, the strong similarity between the overall market shares reported in the original report and the replicated results underscores the robustness and reliability of the findings.
qqq we confirm? "The results are consistent with the view that the primary economic role of reported earnings is not to provide timely new information to the share market. By inference, that role lies elsewhere, for example, in settling debt and compensation contracts and in disciplining prior information,including more timely managerial disclosures of information originating in the firm’s accounting system. The comparatively low “surprise” content of earnings announcements is to be expected, for a variety of reasons. By its nature, accounting earnings is low frequency (quarterly), not discretionary (announced every quarter, independent of whether there is substantial new information to report), and primarily backward-looking. Other information, and hence revision in share price, is comparatively high frequency, frequently discretionary (released only when there is substantial information to report), and both forward-looking and backward-looking."

# Task 2 - Extending the Analysis to 2007–2023 {#sec-task2}
This Section extends the analysis to 2007–2023, a period defined by major economic events such as the 2008 financial crisis and the COVID-19 pandemic, which may have influenced the informativeness of earnings announcements. While the task is relatively straightforward in terms of programming—requiring adjustments to the sample period—it demands careful economic interpretation of observed trends, market shifts, and potential institutional details that may influence the relationship between earnings announcements and stock returns.

@@

Assumptions from @sec-research_design_assumptions1 still hold.

@@

# Task 3 - Cross-Country Replication {#sec-task3}
Task 3 extends the analysis beyond the United States by replicating Figure 1 using 1972-2023 data (if entire data range pull possible) from a non-U.S. market, providing insights into how earnings informativeness varies across different regulatory environments and market structures. This task requires mapping variables from CRSP/Compustat to their equivalents in Worldscope/Datastream, adjusting for differences in accounting standards, market liquidity, and institutional factors. Worldscope (Thomson/Refinitiv) serves as the international alternative of Compustat for earnings announcement data, while Datastream provides stock market data for non-U.S. companies. While the core replication methodology remains the same, careful data handling and economic interpretation are crucial to account for cross-country analysis with Canada.

## Research Design Choices and Assumptions {#sec-research_design_assumptions3}
Since this task employs a generalization and extension approach by applying the same methodology to a country different from original report, it requires adjustments for different databases and market structures. Based on the original paper, I do not gather analyst expectations or estimate earnings deviations for my replication. Instead, I only require stock return data and earnings announcement dates.

Canada as the country of choice was influenced by the alignment with the methodology of @ball_how_2008, where the key factor in the selection was the availability of quarterly earnings reports, which ensures comparability with U.S. data. For Canada, the reporting frequency is majority 4x, with some 2x, similar to the US [@short_short_2025]. While Canada has a smaller market size than the U.S., its regulatory environment, reporting frequency, and institutional settings are highly comparable, making it an optimal choice for the non-U.S. analysis. Additionally, Canada features a developed capital market with high trading volume and liquidity, ensuring that stock price reactions to earnings announcements are observable in the replication. A further advantage is that Canadian firms registered with the U.S. Securities and Exchange Commission (SEC) can report under U.S. GAAP, making their financial disclosures fully compatible with U.S. firms. This reduces methodological differences to the original study.

In addition, I impose the following assumptions apart from those listed in @sec-research_design_assumptions1 to ensure generalization and extension approach due to usage of other databases or where @ball_how_2008 do not provide explicit guidance:

1. To ensure a smooth and conflict-free process, I handle the WSCP/Datastream workflow separately, preventing potential interference with the CRSP/Compustat pull process and improving debugging efficiency. Although a unified pull/prepare/do_analysis file for all databases would be ideal and streamline management, the large-scale nature of the extracted datasets is chosen for modular approach. Therefore, I isolate the data pull for @sec-task3 into `pull_wrds_data-wscp`, thus reducing complicated dependencies across tasks. To maintain structural consistency, I implement same approach for subsequent steps, including data preparation and analysis.

2. Since I do not perform subgroup analysis for December and non-December fiscal year end firms as in Table 2 Panels C-D, `item5350` for FISCAL PERIOD END DATE is not relevant for this task, as Figure 1 applies to the analysis of the entire dataset. So, the sample is all firm-years with available data on the quarterly Worlscope and daily Datastream files.

3. I verified whether Worldscope and Datastream follow the same earnings announcement date convention as Compustat or if they report earnings based on fiscal periods. I decided to use Worldscope items 5901–5904 from historical stock-related data in `wrds_ws_stock`, which represent the actual date when a company publicly reports its earnings for each quarter, where, however, data for these fields is generally not available prior to 1992. [@thomson_financial_worldscope_2007, p.281]. This may cause limited data sample size.

4. This task is a generalization to a non-U.S. setting, meaning methodological differences will certainly arise due to variations in data structure and reporting standards. Worldscope - Stock Data `wrds_ws_stock` table is updated quarterly, while Datastream Daily Stock File `wrds_ds2dsf` is updated weekly, which may introduce timing mismatches when aligning financial fundamentals with stock returns. Given that, future applications should consider database updates and coverage differences between CRSP and Datastream.

5. Similar to @sec-task1, I use returns `ret` variable in Datastream to simplify the process, presented as percentage change. Bid and ask prices are ignored.

6. I ignore the currency effect (CAD or USD) in DS table, as returns are measured in percentages, making them relative rather than absolute values.

## Replication Steps {#sec-replication_steps3}
Same as in @sec-task1 the analysis involves gathering and filtering stock return and earnings announcement data, linking Datastream stock data with WSCP (Worldscope), and estimating the relationship between annual returns and earnings-announcement window returns. The replication process follows a structured workflow of data pull, preparation, and analysis, ensuring consistency with the original study.

**Step 1: Pulling the Data and Managing the Databases**  
I follow the linking approach between Worldscope and Datastream based on @dai_research_2021. The original study uses data from 1972 onward, while Worldscope's `year_` variable only provides data starting from 1980 [@wharton_research_data_services_worldscope_2025]. This constraint may lead to differences in sample coverage. Hence, filter Datastream `marketdate` to 1980 to prevent mismatched observations where stock data exists but earnings do not.

The data pull step retrieves quarterly earnings announcement dates from Worldscope and daily stock returns from Datastream, filtering for Canadian firms from 1980 onward. The datasets are saved separately and later merged using firm-level identifiers. Only common equity (`typecode` = EQ) is included to ensure consistency with CRSP.

**Step 2: Data Preparation**  
If a firm has missing values for any of the Earnings Per Share Report Date fields (`item5901`–`item5904`), it does not meet the criteria set by @ball_how_2008 and was excluded from the sample.

**Step 3: Analysis Implementation and Reproduction of Tables and Figure**

## Results {#sec-results3}
Comparing the replication results with those from @ball_how_2008 (see @fig-fig1_original) reveals both similarities and discrepancies.

qqq Comparing the replication results from Task 1&2 (see @ and @) with those from @ball_how_2008 (see @fig-fig1_original) reveals both similarities and discrepancies.

@@

The study contributes to the ongoing discussion on market efficiency and the economic role of corporate disclosures by providing empirical evidence on how earnings informativeness has evolved over time and across jurisdictions. The cross-country analysis with Canada highlights variations in market responses to earnings announcements, suggesting that institutional factors, market structures, and investor behavior play a role in shaping how financial information is processed. These findings have implications for investors, analysts, and policymakers, as they underscore the importance of regulatory consistency and market transparency in shaping the effectiveness of financial reporting.
